### 2021-08-03 R/remove_climber_sim.Rmd
update remove-climber simulations rmd

---
title: "How Removing a Ranked Climber Changes the Winners"
author: "Hannah Butler"
date: "7/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)

# load data
load("medal_displacement_probabilities.rdata")
load("top3_displace_probs.rdata")
load("medal_displacement_probabilities.rdata")

##### functions #####
assign_ranks <- function(c = 8) {
  # randomly assign ranks in 3 events & compute total
  # ranking in an event is considered to be independent from ranking in another event
  output <- data.frame(climber = sample(LETTERS, c, replace = FALSE)
                       , speed = sample(1:c, c, replace = FALSE)
                       , bould = sample(1:c, c, replace = FALSE)
                       , lead = sample(1:c, c, replace = FALSE)
                       ) %>%
    mutate(total = speed*bould*lead) %>%
    arrange(total) %>% # some questionable tie-breaking code
    mutate(m_rank_speed = ifelse(total %in% total[1:3] , rank(speed[1:3]), 0)
           , m_rank_bould = ifelse(total %in% total[1:3] , rank(bould[1:3]), 0)
           , m_rank_lead = ifelse(total %in% total[1:3] , rank(lead[1:3]), 0)
           , tiebreaker = m_rank_speed + m_rank_bould + m_rank_lead
           ) %>%
    arrange(total
            , tiebreaker
            )

  return(output)
}

drop_climber <- function(results, r = 4) {
  # drop a competitor and recompute rankings and totals
  # r = which rank competitor to drop
  ud_results <- results[-r, ] %>%
    mutate(speed = rank(speed)
           , bould = rank(bould)
           , lead = rank(lead)
           , total = speed*bould*lead
           ) %>%
    arrange(total)%>%
    mutate(m_rank_speed = ifelse(total %in% total[1:3], rank(speed[1:3]), 0)
           , m_rank_bould = ifelse(total %in% total[1:3], rank(bould[1:3]), 0)
           , m_rank_lead = ifelse(total %in% total[1:3], rank(lead[1:3]), 0)
           , tiebreaker = m_rank_speed + m_rank_bould + m_rank_lead
           ) %>%
    arrange(total
            , tiebreaker
            )
  return(ud_results)
}

displace_medalist <- function(nsim = 1000, drop_rank = 4, nc = 8) {
  # simulate nsim rankings
  rank_perms <- lapply(1:nsim, function(x) assign_ranks(c = nc))
  # drop ith competitor and recalculate ranks
  dropout_ranks <- lapply(rank_perms, function(x) drop_climber(x, r = drop_rank))
  # compare 1st-3rd place & calculate % not matched
  pct_displ <- sapply(1:nsim, 
                      function(x) {
                        sum(rank_perms[[x]]$climber[1:3] != dropout_ranks[[x]]$climber[1:3]) > 0
                      }
                      ) %>%
    sum()/nsim

  return(pct_displ)
}

displace_climber <- function(nsim = 1000, drop_rank = 4, nc = 8) {
  # simulate nsim rankings
  rank_perms <- lapply(1:nsim, function(x) assign_ranks(c = nc))
  # drop ith competitor and recalculate ranks
  dropout_ranks <- lapply(rank_perms, function(x) drop_climber(x, r = drop_rank))
  # compare 1st - (r-1)th place & calculate % not matched
  pct_displ <- sapply(1:nsim, 
                      function(x) {
                        sum(rank_perms[[x]]$climber[1:(drop_rank - 1)] != dropout_ranks[[x]]$climber[1:(drop_rank - 1)]) > 0
                      }
                      ) %>%
    sum()/nsim

  return(pct_displ)
}
```

# Intro
A scoring system that multiplies the ranks across events is interesting because it does not possess the property that it is invariant to reordering when a competitor is removed. Examples of this can be found in existing combined climbing event data, where when a climber with a total rank of $r > 1$, higher ranks change order. This has to do with the Independence of Irrelevant Alternatives (IIA) in Voting Systems, which will be discussed later.

We can test this in simulation to see how often we would expect there to be a reordering in competitors when a competitor of a lower rank is removed. In particular, we can estimate the probability that the set of finalists will change if a non-finalist is removed from the rankings, and we can estimate the probability of medalists being displaced if a non-medalist is removed.

## Independence of Irrelevant Alternatives

# Methodology

As it is the relevant upcoming event, we will focus the parameters of the simulation on those of sport climbing 2020 Tokyo Olympics (taking place in August, 2021). Namely, the qualifying round will consist of 40 individuals (20 men, 20 women) and the finals will consist of the 12(16?) (6(8?) men, 6(8?) women) climbers with the highest ranking in the qualifying round. For our simulations, we will naively assume that the ranks in all three events are assigned randomly and independently. (see the section on rank-product statistic which tests this assumption).

# Results

## Probability of Reordering in Higher Ranks
We first look at the rate at which the ranks above that of which is dropped are reordered.
```{r, echo = FALSE}
load("reorder_all_above_probs.rdata")

sum_dat1 <- reorder_all_above %>%
  group_by(dropped_rank) %>%
  rename("Dropped Rank" = dropped_rank) %>%
  summarize("Simulations" = n()
            , "Avg. Probability" = mean(displ_prob)
            , "Median Probability" = median(displ_prob)
            , "St. Deviation" = sd(displ_prob)
            )

knitr::kable(sum_dat1)
```

## Probability of Change in Finalists

A potential event of interest is the replacement of one finalist with another, taking the chance away from the original finalist to win a medal in the finals round. Here we compute the probability that a finalist will be replaced with a non-finalist after removing another non-finalist from the rankings. Below is a table summarizing the results.

```{r, echo=FALSE}
load("drop_qualifier_sims.rdata")

reorder_dat2 <- lapply(9:20
                  , function(x) {
                    data.frame(dropped_rank = x
                               , displ_prob = pct_displace_finalists[[x-8]]
                               )
                    }
                  ) %>%
  do.call(rbind, .)

sum_dat2 <- reorder_dat2 %>%
  group_by(dropped_rank) %>%
  rename("Dropped Rank" = dropped_rank) %>%
  summarize("Simulations" = n()
            , "Avg. Probability" = mean(displ_prob)
            , "Median Probability" = median(displ_prob)
            , "St. Deviation" = sd(displ_prob)
            )

knitr::kable(sum_dat2)

```

as can be seen, the probability is relatively low when the 9th place competitor is removed. My hypothesis is that there is enough separation between the original finalists and the 10th+ ranking competitors in the individual events that there isn't a high chance that the total rankings of the finalists can be overtaken by 10+th place competitors. Note: it might be worth computing the probability of getting a low ranking in any of the individual events, given that a competitor is a finalist.

## Probability that Medalists are Reordered or Changed

The second scenario of interest is if a medalist is displaced by another medalist, or even a non-medalist. Below are the results summarizing the probability of this scenario occurring under uniform random ranking conditions.

```{r, echo = FALSE}
# probability that medalists are displaced (changed or reordered)

displace_probs %>%
  group_by(dropped_rank) %>%
  rename("Dropped Rank" = dropped_rank) %>%
  summarize("Simulations" = n()
            , "Avg. Probability" = mean(displ_prob)
            , "Median Probability" = median(displ_prob)
            , "St. Deviation" = sd(displ_prob)
            ) %>%
  knitr::kable()

```

# prob of displacing medalists as list of competitors grows
```{r, eval=FALSE, include=FALSE}
set.seed(80085)
remove_c4 <- lapply(5:20, function(x) replicate(100, displace_climber(nsim = 100, drop_rank = 4, nc = x)))

remove_c4_df <- lapply(5:20
                  , function(x) {
                    data.frame(n_climbers = x
                               , dropped_rank = 4
                               , displ_prob = remove_c4[[x-4]]
                               )
                    }
                  ) %>%
  do.call(rbind, .)
#save(remove_c4_df, "R/remove_c4_sims.rdata")

remove_c5 <- lapply(5:20, function(x) replicate(100, displace_climber(nsim = 100, drop_rank = 5, nc = x)))

remove_c5_df <- lapply(5:20
                  , function(x) {
                    data.frame(n_climbers = x
                               , dropped_rank = 5
                               , displ_prob = remove_c4[[x-4]]
                               )
                    }
                  ) %>%
  do.call(rbind, .)
#save(remove_c5_df, "R/remove_c5_sims.rdata")
```

### Removing 4th Place
```{r, echo = FALSE}
load("remove_c4_sims.rdata")
remove_c4_df %>%
  ggplot(aes(x = n_climbers, y = displ_prob)) +
  geom_boxplot(aes(group = n_climbers)) +
  theme_minimal() +
  ylab("Probability of Displacing Top 3") +
  xlab("Number of Original Climbers")
```

# Questions: 
- How often would a finalist be replaced
- How often does someone who wasn't a medalist before become a medalist 
- How often do we see a flip (two consecutive climbers switch places), jump (a climber jumps up more than one rank), drop (a climber drops down more than one rank)
- how do the simulations compare to doing the same with actual data?

# Appendix: R Functions

**Randomly generate ranks for 3 events**
```{r, eval=FALSE}
assign_ranks <- function(c = 8) {
  # randomly assign ranks in 3 events & compute total
  # ranking in an event is considered to be independent from ranking in another event
  output <- data.frame(climber = sample(LETTERS, c, replace = FALSE)
                       , speed = sample(1:c, c, replace = FALSE)
                       , bould = sample(1:c, c, replace = FALSE)
                       , lead = sample(1:c, c, replace = FALSE)
                       ) %>%
    mutate(total = speed*bould*lead) %>%
    arrange(total) %>% # some questionable tie-breaking code
    mutate(m_rank_speed = ifelse(total %in% total[1:3] , rank(speed[1:3]), 0)
           , m_rank_bould = ifelse(total %in% total[1:3] , rank(bould[1:3]), 0)
           , m_rank_lead = ifelse(total %in% total[1:3] , rank(lead[1:3]), 0)
           , tiebreaker = m_rank_speed + m_rank_bould + m_rank_lead
           ) %>%
    arrange(total
            , tiebreaker
            )

  return(output)
}
```

**Remove a climber from simulated data and rerank remaining**
```{r, eval=FALSE}
drop_climber <- function(results, r = 4) {
  # drop a competitor and recompute rankings and totals
  # r = which rank competitor to drop
  ud_results <- results[-r, ] %>%
    mutate(speed = rank(speed)
           , bould = rank(bould)
           , lead = rank(lead)
           , total = speed*bould*lead
           ) %>%
    arrange(total)%>%
    mutate(m_rank_speed = ifelse(total %in% total[1:3], rank(speed[1:3]), 0)
           , m_rank_bould = ifelse(total %in% total[1:3], rank(bould[1:3]), 0)
           , m_rank_lead = ifelse(total %in% total[1:3], rank(lead[1:3]), 0)
           , tiebreaker = m_rank_speed + m_rank_bould + m_rank_lead
           ) %>%
    arrange(total
            , tiebreaker
            )
  return(ud_results)
}
```

**Simulate reranking of all climbers above dropped climber**
```{r, eval=FALSE}
displace_climber <- function(nsim = 1000, drop_rank = 4, nc = 8) {
  # simulate nsim rankings
  rank_perms <- lapply(1:nsim, function(x) assign_ranks(c = nc))
  # drop ith competitor and recalculate ranks
  dropout_ranks <- lapply(rank_perms, function(x) drop_climber(x, r = drop_rank))
  # compare 1st - (r-1)th place & calculate % not matched
  pct_displ <- sapply(1:nsim, 
                      function(x) {
                        sum(rank_perms[[x]]$climber[1:(drop_rank - 1)] != dropout_ranks[[x]]$climber[1:(drop_rank - 1)]) > 0
                      }
                      ) %>%
    sum()/nsim

  return(pct_displ)
}
```

**Simulate reranking of medalists when a non-medalist is dropped**
```{r, eval=FALSE}
displace_medalist <- function(nsim = 1000, drop_rank = 4, nc = 8) {
  # simulate nsim rankings
  rank_perms <- lapply(1:nsim, function(x) assign_ranks(nc))
  # drop ith competitor and recalculate ranks
  dropout_ranks <- lapply(rank_perms, function(x) drop_climber(x, r = drop_rank))
  # compare 1st-3rd place & calculate % not matched
  pct_displ <- sapply(1:nsim, 
                      function(x) {
                        sum(rank_perms[[x]]$climber[1:3] != dropout_ranks[[x]]$climber[1:3]) > 0
                      }
                      ) %>%
    sum()/nsim

  return(pct_displ)
}
```

**Simulate replacement of finalists when a non-finalist is dropped** 
*permutations do not count*
```{r, eval=FALSE}
displace_finalist <- function(nsim = 100, drop_rank = 9) {
  # simulate nsim rankings
  rank_perms <- lapply(1:nsim, function(x) assign_ranks(c = 20))
  # drop ith competitor and recalculate ranks
  dropout_ranks <- lapply(rank_perms, function(x) drop_climber(x, r = drop_rank))
  # compare 1st-3rd place & calculate % not matched
  pct_displ <- sapply(1:nsim, 
                      function(x) {
                        sum(rank_perms[[x]]$climber[1:8] %in% dropout_ranks[[x]]$climber[1:8]) < 8
                      }
  ) %>%
    sum()/nsim

  return(pct_displ)
}
```