### 2021-08-13 R/climber_swaps.Rmd
results after swapping of first/second place in each event

---
title: "swaps"
author: "Hannah Butler"
date: "8/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
load(file = "2020_olympics.rdata")
```

### Men's Olympic Finals

```{r, echo = FALSE}
mens_finals_events <- mens_finals %>%
  select(climber, og_speed = speed, og_bould = bouldering, og_lead = lead)

mens_swaps12 <- mens_finals_events %>%
  mutate(swap_speed = ifelse(climber == "Alberto Ginés López", 2, og_speed)
         , swap_speed = ifelse(climber == "Tomoa Narasaki", 1, swap_speed)
         , swap_bould = ifelse(climber == "Nathaniel Coleman", 2, og_bould)
         , swap_bould = ifelse(climber == "Mickaël Mawem", 1, swap_bould)
         , swap_lead = ifelse(climber == "Jakob Schubert", 2, og_lead)
         , swap_lead = ifelse(climber == "Adam Ondra", 1, swap_lead)
         ) %>%
  pivot_longer(og_speed:swap_lead) %>%
  rename(swap = name) %>%
  mutate(event = sub(".*_", "", swap)) %>%
  mutate(swap = ifelse(substr(swap, 1,2) == "og", "original", sub(".*_", "", swap))) %>%
  pivot_wider(names_from = event, values_from = value) %>%
  group_by(climber) %>%
  mutate(speed = ifelse(is.na(speed), speed[swap == "original"], speed)
         , bould = ifelse(is.na(bould), bould[swap == "original"], bould)
         , lead = ifelse(is.na(lead), lead[swap == "original"], lead)
         ) %>%
  ungroup() %>%
  group_by(swap) %>%
  mutate(total = speed*bould*lead
         , total_rank = rank(total)
         ) %>%
  arrange(swap, total_rank)

mens_swaps12 %>% 
  filter(swap == "original") %>%
  knitr::kable()

mens_swaps12 %>% 
  filter(swap == "speed") %>%
  knitr::kable()

mens_swaps12 %>% 
  filter(swap == "bould") %>%
  knitr::kable()

mens_swaps12 %>% 
  filter(swap == "lead") %>%
  knitr::kable()
```

### Women's Olympic Finals

```{r, echo = FALSE}
womens_finals_events <- womens_finals %>%
  select(climber, og_speed = speed, og_bould = bouldering, og_lead = lead)

womens_swaps12 <- womens_finals_events %>%
  mutate(swap_speed = ifelse(climber == "Aleksandra Mirosław", 2, og_speed)
         , swap_speed = ifelse(climber == "Anouck Jaubert", 1, swap_speed)
         , swap_bould = ifelse(climber == "Janja Garnbret", 2, og_bould)
         , swap_bould = ifelse(climber == "Brooke Raboutou", 1, swap_bould)
         , swap_lead = ifelse(climber == "Janja Garnbret", 2, og_lead)
         , swap_lead = ifelse(climber == "Seo Chae-hyun", 1, swap_lead)
         ) %>%
  pivot_longer(og_speed:swap_lead) %>%
  rename(swap = name) %>%
  mutate(event = sub(".*_", "", swap)) %>%
  mutate(swap = ifelse(substr(swap, 1,2) == "og", "original", sub(".*_", "", swap))) %>%
  pivot_wider(names_from = event, values_from = value) %>%
  group_by(climber) %>%
  mutate(speed = ifelse(is.na(speed), speed[swap == "original"], speed)
         , bould = ifelse(is.na(bould), bould[swap == "original"], bould)
         , lead = ifelse(is.na(lead), lead[swap == "original"], lead)
         ) %>%
  ungroup() %>%
  group_by(swap) %>%
  mutate(total = speed*bould*lead
         , total_rank = rank(total)
         ) %>%
  arrange(swap, total_rank)

womens_swaps12 %>% 
  filter(swap == "original") %>%
  knitr::kable()

womens_swaps12 %>% 
  filter(swap == "speed") %>%
  knitr::kable()

womens_swaps12 %>% 
  filter(swap == "bould") %>%
  knitr::kable()

womens_swaps12 %>% 
  filter(swap == "lead") %>%
  knitr::kable()
```

### How unstable are the medalists in their positions?